Review: The Prevalence of Adaptive Introgression in Crop Histories

### 2018-06-26 paper.tex
edits to domestication questions

\subsection*{What is the progenitor of a crop?}
Depending on the extent of post-domestication gene flow with new wild relatives, identification of a crop's progenitor can be complicated or confounded entirely.
Introgression between a crop and newly-encountered taxa decreases divergence of the crop from these donors.
This signal could be mistaken for origin rather than gene flow.
For example, when determining a single origin of maize from \emph{parviglumis}, Matsuoka and colleagues \cite{matsuoka2002single} identified a paradox: while \emph{parviglumis} is found exclusively in the lowlands of southwest Mexico, maize with allele frequencies most similar to \emph{parviglumis} was found in the highlands of the Mexican Central Plateau.
Several years later, van Heerwaarden \emph{et al.} \cite{vanHeerwaarden2011} resolved the paradox by determining that widespread introgression in the highlands from \emph{mexicana}, which is closely related to \emph{parviglumis}, has caused maize from this region to appear ancestral.
Similarly, extensive post-domestication adaptive introgression from  potato wild relatives long obscured this crop's origin.
Recent work has shown that, following the original domestication event of \emph{Solanum tuberosum} in the central Andes, potato received introgression from as many as four additional species during colonization of the highest elevations of the Andes and the lowlands of the Chilean coast \cite{Spooner2014, Gavrilenko2013, Hardigan2017}.

Beyond confounding detection of progenitor taxa, extensive introgression may necessitate a reevaluation of crop origins.
In cases like maize and potato it is important to recognize the substantial contributions of introgressing taxa to the genetic base of modern crops.
Broad recognition of the role these wild relatives have played in crop adaptation could further their use in breeding and elevate their conservation status.

\subsection*{When was a crop domesticated?}
Estimates of the timing of initial domestication are often based on levels of sequence divergence between a crop and populations of its presumed progenitor (\emph{e.g.}, \cite{matsuoka2002single, molina2011molecular}).
In highly introgressed domesticates, these estimates will be based on comparison of both crop and introgressant haplotypes to those of the presumed progenitor.
In such cases, divergence time is a mixture of time since domestication and time since split of the progenitor and the introgressing taxon.
This phenomenon, in combination with divergence of crop samples from true ancestral crop populations, ongoing evolution of crop progenitors, and problems with assuming evolution under a molecular clock \cite{Zeder2006}, may help explain discrepancies between domestication dates based on genetic and archaeological data.
More accurate estimates of the timing of domestication may be obtained from genetic data by excluding loci that show signatures of introgression or by explicitly including estimates of introgression when modeling a crop's demographic history.

\subsection*{How was genome-wide diversity shaped by domestication?}

Measurement of the strength of the initial domestication bottleneck may also be impacted by adaptive introgression during the spread of crops.
Crop wild relatives have distinct demographies when compared to domesticates and may therefore have contrasting effective population sizes ($N_e$).
The influence of wild relative introgression on estimates of the domestication bottleneck will depend on a number of factors including the magnitude of gene flow, the $N_e$ of the introgressing taxon, and the strength of selection on haplotypes following introgression.
For example, substantial introgression at neutral loci from a wild taxon with a historically higher $N_e$ will lead to underestimates of the overall strength of the initial domestication bottleneck.

\subsection*{What candidate genes were targeted by selection during domestication?}
Loci targeted by selection during domestication can be identified through so-called ``bottom-up'' approaches based on population genetic signatures \cite{Ross-Ibarra2007}.
Ideally, candidate loci will be identified by first constructing a demographic model representing the history of the domesticate.
In this approach, polymorphism data from neutral loci are fit to potential models of a crop's demography and then statistical tests of selection are used to identify candidate domestication genes under the most likely model.
Due to the difficulty of this approach and the uncertainty associated with any given demography, many studies identify domestication loci using a strict outlier approach in which loci showing, for example, the greatest reduction in nucleotide diversity or the highest allele frequency differentiation in the domesticate relative to the wild progenitor are identified as candidates.
Introgression during crop expansion may influence candidate gene detection using both demographic-modeling and strict-outlier approaches.
For example, \emph{mexicana} introgression into maize described above accounts for approximately 20\% of the genome of maize in the highlands of Mexico \cite{vanHeerwaarden2011}.
Takuno and co-authors \cite{Takuno2015} have shown that a demographic model incorporating this introgression is a significantly better fit to empirical data than a model lacking introgression.
Failure to account for introgression in maize would therefore compromise domestication candidate detection, particularly if a study contained maize samples from the Mexican highlands.
Likewise, introgression that increased nucleotide diversity in the domesticate or decreased differentiation at domestication loci would confound a strict outlier approach.
However, previous work, also in maize, has shown that known domestication loci are particularly resistant to introgression \cite{Hufford2013}, likely due to ongoing selection favoring the domesticated phenotype.

\section*{Evaluating crop adaptation through introgression:}

>>>>

\subsection*{Where and from what taxon did a crop originate?}
Depending on the extent of post-domestication gene flow with new wild relatives, identification of a crop's origin may be complicated or confounded entirely.
Introgression between a crop and newly-encountered taxa decreases divergence of the crop from these donors.
This signal could be mistaken for origin rather than gene flow.
For example, when determining a single origin of maize from \emph{parviglumis}, Matsuoka and colleagues \cite{matsuoka2002single} identified a paradox: while \emph{parviglumis} is found exclusively in the lowlands of southwest Mexico, maize with allele frequencies most similar to \emph{parviglumis} was found in the highlands of the Mexican Central Plateau.
Several years later, van Heerwaarden \emph{et al.} \cite{vanHeerwaarden2011} resolved the paradox by determining that widespread introgression in the highlands from \emph{mexicana}, which is closely related to \emph{parviglumis}, has caused maize from this region to appear ancestral.
Similarly, as described above, extensive post-domestication adaptive introgression from  potato wild relatives long obscured this crop's origin.

Beyond confounding detection of progenitor taxa, extensive introgression may necessitate a more nuanced view of crop origins.
In cases like maize and potato it is important to recognize the substantial contributions of introgressing taxa to modern crops.
While these crops may have originated from a single species or subspecies, the crops as we know them today have a broader genetic base.

\subsection*{When was a crop domesticated?}
Estimates of the timing of initial domestication are often based on levels of sequence divergence between a crop and populations of its presumed progenitor (\emph{e.g.}, \cite{matsuoka2002single, molina2011molecular}).
In highly introgressed domesticates, these estimates will be based on comparison of both crop and introgressant haplotypes to those of the presumed progenitor.
In such cases, divergence time is a mixture of time since domestication and time since split of the progenitor and the introgressing taxa.
This phenomenon, in combination with divergence of modern crop samples from true ancestral crop populations, ongoing evolution of crop progenitors, and problems with assuming evolution under a molecular clock \cite{Zeder2006}, may help explain discrepancies between domestication dates based on genetic and archaeological data.
More accurate estimates of the timing of domestication may be obtained from genetic data by excluding loci that show signatures of introgression or by explicitly including estimates of introgression when modeling a crop's demographic history.

\subsection*{How has genome-wide diversity been shaped by domestication?}

Measurement of the strength of the initial domestication bottleneck may also be impacted by adaptive introgression during the spread of crops.
Crop wild relatives have distinct demographies when compared to domesticates and may therefore have contrasting effective population sizes ($N_e$).
The influence of wild relative introgression on estimates of the domestication bottleneck will depend on a number of factors including the magnitude of gene flow, the $N_e$ of the introgressing taxon, and the strength of selection on haplotypes following introgression.
For example, substantial introgression from a wild taxon with a historically higher $N_e$ will lead to underestimates of the overall strength of the initial domestication bottleneck.

\subsection*{What candidate genes were targeted by selection during domestication?}
Loci targeted by selection during domestication can be identified through so-called ``bottom-up'' approaches based on population genetic signatures \cite{Ross-Ibarra2007}.
Ideally, candidate loci will be identified by first constructing a demographic model representing the history of the domesticate.
In this approach, polymorphism data from neutral loci are fit to potential models of a crop's demography and then statistical tests of selection are used to identify candidate domestication genes under the most likely model.
Due to the uncertainty associated with any given demography, many studies identify domestication loci using a strict outlier approach in which loci showing, for example, the greatest reduction in nucleotide diversity or the highest allele frequency differentiation in the domesticate relative to the wild progenitor are identified as candidates.
Introgression during crop expansion may influence candidate gene detection using both demographic-modeling and strict-outlier approaches.
For example, \emph{mexicana} introgression into maize described above accounts for approximately 20\% of the genome of maize in the highlands of Mexico \cite{vanHeerwaarden2011}.
Takuno and co-authors \cite{Takuno2015} have shown that a demographic model incorporating this introgression is a significantly better fit to empirical data than a model lacking introgression.
Failure to account for introgression in maize would therefore compromise domestication candidate detection, particularly if a study contained maize samples from the Mexican highlands.
Likewise, introgression that increased nucleotide diversity in the domesticate or decreased differentiation at domestication loci would confound a strict outlier approach.
However, previous work, also in maize, has shown that known domestication loci are particularly resistant to introgression \cite{Hufford2013}, likely due to ongoing selection favoring the domesticated phenotype.
\vspace{5mm}

In summary, should post-domestication gene flow with wild relatives be pervasive during crop histories, investigations seeking to unravel fundamental questions of crops' initial domestication must take this into account in order to accurately estimate parameters of interest.

\section*{Investigating crop adaptation through introgression:}
